---
title: "Homework 4"
author: "Math 445, Spring 2017"
date: "Due Wednesday, May 3 by 4:30 p.m."
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Instructions

- Please complete problems that require the use of R in R markdown. Theoretical problems may be handwritten. **Please knit your .Rmd file to either Word or PDF and print it for submission** with the remainder of the problems.

- All assignments should be stapled.

- Remember that the textbook data sets are contained in the `resampledata` R package. 

- When completing a hypothesis test, be sure to outline all of the steps in your solution.

## Assignment

Complete the following exercises from **Chapter 6** of the textbook. 

**Exercise 11**

**Exercise 16**

*Notes on part b:* If you use R to run the goodness-of-fit test, include the results (the value of the test statistic and associated p-value) and the R code used in your solution.

*Notes on part c:* To superimpose the Gamma($r$, $\lambda$) PDF on a histogram (remember to change the scale of the y-axis to the density scale!), use the below layer:

```{r eval=FALSE}
stat_function(fun = dgamma, geom = "line", args = list(shape = r, rate = lambda))
```

where `r` and `lambda` are the estimates of $r$ and $\lambda$, respectively.

To superimpose the Gamma($r$, $\lambda$) CDF on a plot of the ECDF, use the below layer:

```{r eval=FALSE}
stat_function(fun = pgamma, geom = "line", args = list(shape = r, rate = lambda))
```


You can also experiment with the  `color` and `linetype` arguments to make your plots easier to read.

**Exercise 19**

**Exercise 29**