---
title: "Numeric ML estimate in R"
author: "Math 445, Spring 2017"
output: html_document
---

This R markdown file provides my code for the Boston Storm example of using `optim` to find ML estimates in R for the gamma distribution using `optim`.

### Loading our usual packages and the data

```{r}
library(ggplot2)
storm <- read.csv("BostonStorms2013.csv")
```

### Plotting the distribution of precipitation

```{r}
ggplot(data = storm, mapping = aes(x = precip)) + 
  geom_histogram(colour = "gray20")

ggplot(data = storm, mapping = aes(x = precip)) + 
  geom_density()
```


### Calculating the MoM estimates

```{r}
n <- length(storm$precip) 
xbar <- mean(storm$precip) 
s2 <- var(storm$precip) 
sigma2_hat <- s2 * (n - 1) / n 

lambda_mom <- xbar / sigma2_hat; lambda_mom
alpha_mom <- xbar^2 / sigma2_hat; alpha_mom
```


### Writing the log-likelihood function

```{r}
gamma_loglik <- function(theta, x) {
  n <- length(x)
  alpha <- theta[1]
  lambda <- theta[2]
  RES <- n * (alpha * log(lambda) - log(gamma(alpha)) + 
                (alpha - 1) * mean(log(x)) - lambda * mean(x)) 
  return(RES)
}
```


### Using optim to maximize the log-likelihood

```{r}
theta <- optim(par = c(alpha_mom, lambda_mom), fn = gamma_loglik, 
               control = list(fnscale = -1), x = storm$precip)
theta
```


### Alternative way to write the log-likelihood (RECOMMENDED)

When possible, use the built in mass/density functions for your log likelihoods!

```{r}
gamma_loglik_alt <- function(theta, x) {
  sum(dgamma(x, shape = theta[1], rate = theta[2], log = TRUE))
}
```


### Using optim to maximize the log-likelihood

The results are the same here.

```{r}
theta2 <- optim(par = c(alpha_mom, lambda_mom), fn = gamma_loglik_alt, 
                control = list(fnscale = -1), x = storm$precip)
theta2
```

